One-Particle Density Matrix Polarization Susceptibility Tensors

### 2018-07-27 papers/1_dmatpol/paper-r3.tex
Equation in the Appendix B updated

The gradient vector ${\bf g}$ and Hessian matrix ${\bf H}$ 
are build from blocks associated with the type of parameters, i.e.,
%
\begin{equation}\label{e:Newton.gH}
 {\bf g} = 
\begin{pmatrix}
{\bf g}^{[1]} \\ 
{\bf g}^{[2]} 
\end{pmatrix} ,\quad
 {\bf H} = 
\begin{pmatrix}
{\bf H}^{[11]} & {\bf H}^{[12]}  \\ 
{\bf H}^{[21]} & {\bf H}^{[22]}  
\end{pmatrix} \;.
\end{equation}
%
The gradient element of the $r$th block and Hessian element of the $(rs)$th block read
%
\begin{subequations}
 \begin{align}
  g^{[r ]}_{kv}    &\equiv \frac{\partial   Z}{\partial S_{kv}^{[r]}} 
     =-2\sum_N \overline{\delta D}^{(N)}
               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{kv}^{[r]}} \;,\\
  H^{[rs]}_{kv,lw} &\equiv \frac{\partial^2 Z}{\partial S_{kv}^{[r]} \partial S_{lw}^{[s]}}  
     = 2\sum_N 
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{kv}^{[r]}}
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S_{lw}^{[s]}} \;.
 \end{align}
\end{subequations}
%
Note that the second derivatives of $\delta D^{(N)}$ vanish
due to the linearity with respect to the adjustable parameters.
Thus, the explicit formulae for the gradient are
%
\begin{subequations}
 \begin{align}
  g^{[1]}_{kv} &=-2\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \;,\\
  g^{[2]}_{kv} &=-4\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \widetilde{F}^{(N)}_k \;,\\
  g^{[3]}_{kv} &=-4\sum_N \overline{\delta D}^{(N)} \widetilde{F}^{(N)}_{kv} \;.
 \end{align}
\end{subequations}
%
The Hessian subsequently follows to be
%
\begin{subequations}
 \begin{align}
  H^{[11]}_{kv,lw} &= 2\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \;,\\
  H^{[22]}_{kv,lw} &= 8\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \widetilde{F}^{(N)}_k \widetilde{F}^{(N)}_l \;,\\
  H^{[33]}_{kv,lw} &= 8\sum_N \widetilde{F}^{(N)}_{kv} \widetilde{F}^{(N)}_{lw} \;,\\
  H^{[12]}_{kv,lw} &= 4\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \widetilde{F}^{(N)}_l \;,\\
  H^{[13]}_{kv,lw} &= 4\sum_N F^{(N)}_{kv} \widetilde{F}^{(N)}_{lw} \;,\\
  H^{[23]}_{kv,lw} &= 8\sum_N F^{(N)}_{kv} \widetilde{F}^{(N)}_k \widetilde{F}^{(N)}_{lw} \;.
 \end{align}
\end{subequations}
%
In the above equations, the auxiliary quantities $\widetilde{F}^{(N)}_k$ 
and $\widetilde{F}^{(N)}_{kv}$ are defined as
%
\begin{subequations}
 \begin{align}
  \widetilde{F}^{(N)}_k    &\equiv \sum_w F^{(N)}_{kw} \;,\\
  \widetilde{F}^{(N)}_{kv} &\equiv \sum_w F^{(N)}_{kw,kv} \;.
 \end{align}
\end{subequations}
%
Note that due to the symmetry of the Hessian matrix, the blocks $21$, $31$ and $32$
are transposes of the blocks $12$, $13$ and $23$, respectively. The composite index
can be defined as $kv \equiv 3k+v$, for example.
\bibliography{references}

>>>>

The gradient vector ${\bf g}$ and Hessian matrix ${\bf H}$ 
are build from blocks associated with a particular type of parameters, i.e.,
%
\begin{equation}\label{e:Newton.gH}
 {\bf g} = 
\begin{pmatrix}
{\bf g}^{[1]} \\ 
{\bf g}^{[2]} 
\end{pmatrix} ,\quad
 {\bf H} = 
\begin{pmatrix}
{\bf H}^{[11]} & {\bf H}^{[12]}  \\ 
{\bf H}^{[21]} & {\bf H}^{[22]}  
\end{pmatrix} \;.
\end{equation}
%
where the block indices 1 and 2 correspond to the first\hyp{} and second\hyp{}order susceptibilities, respectively.
Note that the second derivatives of $\delta D^{(N)}$ 
with respect to the adjustable parameters vanish
due to the linear functional form of Eq.~\eqref{e:final-model.General.Parameters}.
Thus, the gradient element of the $r$th block and Hessian element of the $(rs)$th block read
%
\begin{subequations}
 \begin{align}
  g^{[r ]}    &\equiv \frac{\partial   Z}{\partial S^{[r]}} 
     =-2\sum_N \overline{\delta D}^{(N)}
               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[r]}} \;,\\
  H^{[rs]} &\equiv \frac{\partial^2 Z}{\partial S^{[r]} \partial S^{[s]}}  
     = 2\sum_N 
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[r]}}
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial S^{[s]}} \;.
 \end{align}
\end{subequations}
%
The explicit formulae for the gradient are
%
\begin{subequations}
 \begin{align}
  g^{[1]}_{kv} &=-2\sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} \;,\\
  g^{[2]}_{kvw} &=-2r_{vw} \sum_N \overline{\delta D}^{(N)} F^{(N)}_{kv} F^{(N)}_{kw} \;.
 \end{align}
\end{subequations}
%
The Hessian subsequently follows to be
%
\begin{subequations}
 \begin{align}
  H^{[11]}_{kv,lw} &= 2\sum_N F^{(N)}_{kv} F^{(N)}_{lw} \;,\\
  H^{[12]}_{kv,lv'w'} &= 2r_{v'w'} \sum_N F^{(N)}_{kv} F^{(N)}_{lv'} F^{(N)}_{lw'}  \;,\\
  H^{[22]}_{kvw,lv'w'} &= 2r_{vw} r_{v'w'} \sum_N F^{(N)}_{kv} F^{(N)}_{kw} F^{(N)}_{lv'} F^{(N)}_{lw'} \;.
 \end{align}
\end{subequations}
%
Note that due to the symmetry of the Hessian matrix, the block $21$
is a transpose of the block $12$. 
The composite indices $kv$ and $kvw$ are constructed from the distributed site index $k$
and the appropriate symmetry\hyp{}corrected Cartesian component 
of a particular DMS tensor ($v$ for first\hyp{}order,
and $vw$, $w<v$ for second\hyp{}order susceptibility tensor, respectively).
\bibliography{references}
